---
title: "IL6_adhd"
author: "Anton Glasberg (edited by Love Ahnström)"
date: "2024-02-27"
output_file: README.md
output: md_document
---

Setup knitr, install required packages. 

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

if(!require(metafor)){
    install.packages("metafor")
    library(metafor)
}

if(!require(here)){
    install.packages("here")
    library(here)
}


```

rma.mv for meta-analyses, raw values, (rma.mv --> Leffa is the same patients + control at two different time points)

```{r read_studies}
df <- read.csv(here("data", "studier.csv"))
df <- escalc(
                    measure = "SMD",
                    data=df,
                  	m1i = meanIL6_adhd,
                  	sd1i = sdIL6_adhd,
                  	n1i = n_adhd,
                  	m2i = meanIL6_control,
                  	sd2i = sdIL6_control,
                  	n2i = n_control,
                    append = TRUE
                    )

 
result <- rma.mv(yi, vi, random = ~ 1 | StudyNumber/Group, data = df)
forest(result, slab = df$Studies, main = "Expanded meta-analysis")
result

```

Higgins method 1 for transforming values into log-values as IL-6 is not normally distributed."

```{r}

higgins_1_mean <- function(mean, sd) {
  return(log(mean) - 0.5 * log((sd^2/mean^2) + 1))
}

higgins_1_sd <- function(mean, sd) {
  return(sqrt(log(sd^2/mean^2 + 1)))
}

```


Meta-analysis of skewed data: Combining results reported on log-transformed or raw scales (nih.gov)

```{r}

df <- read.csv(here("data", "studier.csv"))

df$logtransformmean_adhd <- higgins_1_mean(df$meanIL6_adhd, df$sdIL6_adhd)
df$logtransformSD_adhd <- higgins_1_sd(df$meanIL6_adhd, df$sdIL6_adhd)
df$logtransformmean_control <- higgins_1_mean(df$meanIL6_control, df$sdIL6_control)
df$logtransformSD_control <- higgins_1_sd(df$meanIL6_control, df$sdIL6_control)

df <- escalc(
              measure = "SMD",
             	m1i = logtransformmean_adhd,
             	sd1i = logtransformSD_adhd,
             	n1i = n_adhd,
             	m2i = logtransformmean_control,
             	sd2i = logtransformSD_control,
             	n2i = n_control,
             	data = df,
             	append = TRUE
              )
 
logresult <- rma.mv(yi, vi, random = ~ 1 | StudyNumber/Group, data = df)
forest(logresult, slab = df$Studies, main = "log transformed original studies")

```

Combination method: Higgins method 1 and extracted logged values from Leffa reporting logged values. The data in the file 'logexpand' comes from the forest plot in the previous code using method 1 from Higgins. However, the data in Leffa is manually entered into 'logexpand', so all rows except Leffa18 and Leffa22 should have the same estimate in the forest plot.

```{r}

logexpand <- read.csv(here("data", "logexpand.csv"))
logexpand$se <- (logexpand$upper - logexpand$lower) / (qnorm(0.975)*2) # CI --> SE
logexpand$V <- logexpand$se^2
resultlogexpand <- rma.mv(V = V, yi = logexpand$hedgesg, random = ~ 1 | StudyNumber/Group, data = logexpand)
forest(resultlogexpand, slab = logexpand$AStudies, main = "Extended meta-analysis: Log transformation using Higgins except Leffa")

```


Pooling subgroup analysis of studies where individuals with ADHD were unmedicated, coded as a 'standard meta-analysis'

```{r}
df <- read.csv(here("data", "unmedicated.csv"))
df <- escalc(
                      measure = "SMD", 
                      m1i = meanIL6_adhd, 
                      sd1i = sdIL6_adhd, 
                      n1i = n_adhd, 
                      m2i = meanIL6_control, 
                      sd2i = sdIL6_control, 
                      n2i = n_control,
                      data = df,
                      append = T
                      )

result <- rma(yi, vi, data = df)
forest(result, slab = df$Studies, main = "Expanded meta-analysis")
result

```

meta-regressions

```{r}
df <- read.csv(here("data", "studier.csv"))

df$logtransformmean_adhd <- higgins_1_mean(df$meanIL6_adhd, df$sdIL6_adhd)
df$logtransformSD_adhd <- higgins_1_sd(df$meanIL6_adhd, df$sdIL6_adhd)
df$logtransformmean_control <- higgins_1_mean(df$meanIL6_control, df$sdIL6_control)
df$logtransformSD_control <- higgins_1_sd(df$meanIL6_control, df$sdIL6_control)


df <- escalc(measure = "SMD",
             	m1i = logtransformmean_adhd,
             	sd1i = logtransformSD_adhd,
             	n1i = n_adhd,
             	m2i = logtransformmean_control,
             	sd2i = logtransformSD_control,
             	n2i = n_control,
             	data = df
             	)

SmdAge <- escalc(
              measure="SMD", 
              m1i=Age_ADHD_mean, 
              m2i=AgeHealthymean,
             	sd1i=Age_ADHD_SD, 
              sd2i=Age_Healthy_SD,
             	n1i=n_adhd, 
              n2i=n_control,
             	data=df
              )
 
regression <- rma(
                yi, 
                vi,
              	mods = ~ SmdAge$yi, 
                SmdAge$vi, # Formula might be wrong, please double check. This argument does nothing.
              	data=df,
              	method="REML")
summary(regression)
regplot(regression)
```

sex meta-regression

```{r}

procentmen <- df$ADHDprocentmales - df$Healthyprocentmales
regression <- rma(
                yi, 
                vi,
              	mods = ~ procentmen,
              	data=df, method="REML")
 
summary(regression)
regplot(regression)


```

Nos-score metaregression

```{r}
 
regression <- rma(
                yi, 
                vi,
              	mods = ~ df$NOS,
              	data=df, method="REML")
 
summary(regression)
regplot(regression)

```

